\chapter{Introduction}
\label{chap:chapter1}

Just some ideas for the thesis:

Chapter 1: Introduction
in the introduction i should talk about the role of x ray detectors in different sectors, such as astrophysics, medical imaging and also study of matter properties.

Chapter 2: Ionization detectors (sects: physical processes, gas detectors, semiconductors)
Then there should be a chapter where i discuss how ionization detectors work in general, the physical principles and their characteristics. Then, i should go into the details of gas detectors e semi-conductors.

Chapter 3: Experimental Setup or something similar (sects: XPOL-III chip, ASIX setup, muGPD setup and working environments)
i should also discuss the XPOL chip used in the detectors along with the experimental setup (such as the gembox for mugpd or the x-ray tube test for asix). Obviously this is the perfect place to put some images taken during the thesis.

Chapter 4: ASIX (sects: noise calibration, gain calibration, description of various methods of reconstruction with MC, energy reconstruction with MC, comparison with data, with huttner and slanted edge)
this is the main section where i should describe my work. It is also important to describe the MC software that is used for some results. Here i have to talk about position reconstruction, 

Chapter 5: muGPD (sects: detector characterization, study of gain with voltage, energy resolution etc...)
This is maybe the most difficult section, becuse i haven't done much in reality, just a few plots.

\fix{This is a possible structure of the introduction chapter. The idea is to produce a sort of long abstract a few pages long:}

1.1 Context and Motivation (1-2 pagine): Inquadra il problema generale. Parli dell'imaging a raggi X, dell'importanza di lavorare a energie intorno ai 10 keV (ad esempio per sincrotroni, astrofisica, o imaging biomedico ad alto contrasto) e del perché servono rivelatori digitali ad alta risoluzione.

1.2 The Open Problem: Resolution and Charge Sharing (1-2 pagine):
Introduci il nodo cruciale della tua tesi. Spieghi che quando la nuvola si allarga per diffusione trasversa, la carica si divide tra i pixel (charge sharing). Questo fenomeno da un lato degrada l'immagine, ma dall'altro (ed è qui che entri in gioco tu) può essere modellizzato e sfruttato per spingere la risoluzione spaziale (MTF) a livello sub-pixel.

1.3 Thesis Objectives and Structure (1 pagina):
Dichiari esplicitamente l'obiettivo del tuo lavoro ("In questo lavoro di tesi si presenta la caratterizzazione/modellizzazione di...") e chiudi il capitolo con la classica "mappa" della tesi: “Il Capitolo 2 tratterà la fisica del trasporto di carica, il Capitolo 3 descriverà l'apparato sperimentale